---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Loading and preprocessing the data

Load libraries

```{r}
library(knitr)
opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
```

Load the data 
```{r}
data <- read.csv("activity.csv", header = TRUE, sep =",", colClasses = c("numeric", "character","integer"))
```
Process/transform the data
```{r}
data$date <- ymd(data$date)
```
Check the data with str() and head():
```{r}
str(data)
head(data)
```

# What is mean total number of steps taken per day?

Calculate the total number of steps per day using dplyr and group by date:
```{r}
steps <- data %>%
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        summarize(steps = sum(steps)) %>%
        print
```
1. Make a histogram
```{r}
ggplot(steps, aes(x = steps)) +
        geom_histogram(fill = "firebrick", binwidth = 1000) +
        labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```


2. Calculate the mean and median of the total number of steps taken per day:
```{r}
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)
mean_steps
median_steps
```

# What is the average daily activity pattern?

Calculate the average number of steps taken in each 5-minute interval per day using dplyr and group by interval:
```{r}
interval <- data %>%
        filter(!is.na(steps)) %>%
        group_by(interval) %>%
        summarize(steps = mean(steps))
```
1. making the time series of the 5-minute interval and average steps taken:
```{r}
ggplot(interval, aes(x=interval, y=steps)) +
        geom_line(color = "firebrick")
```

2. find out the maximum steps
```{r}
interval[which.max(interval$steps),]
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data$steps))
data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]

sum(is.na(data_full$steps))

steps_full <- data_full %>%
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        summarize(steps = sum(steps)) %>%
        print

ggplot(steps_full, aes(x = steps)) +
        geom_histogram(fill = "firebrick", binwidth = 1000) +
        labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")

mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)
mean_steps_full
median_steps_full
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}

data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)

interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)

```

